---
title: "assignment_2"
author: "Timothy Roy"
date: '2023-03-26'
output: html_document
---
```{r, echo=FALSE, include=FALSE}
library(dplyr)
library(arules)
library(skimr)
library(plyr)
library(RColorBrewer)
library(arulesViz)
library(imbalance)

```
### Task 1 - Product Orders


# Task 1: Product Orders

## a. Data Preparation

### a.1 Familiarize yourself with dataset for product orders



```{r}
store_data <- read.csv("store.csv")
head(store_data)
```

* This data shows information regarding customer such as city, country, product purchase, their quantity and sales and other order related information.

### a.2 Document and eliminate any data quality issues that you may find (if you find any), i.e., perform the necessary data cleaning.

```{r}
store_data %>% skim()
```


* The data set has 51,290 rows of data which has 25 columns. 13 of columns are character and 12 is numeric.
* Among numeric columns, only postal code has 41,296 missing values which would not be important for association mining. 
* The data set has a minimum profit of -6,599.978 which is a negative number. However, profit's value can be negative. 
* The minimum sales value is surprisingly low (0.444)! Not sure what to do about it.


### a.3 Transforming the data 
In order to run association algorithm among products, `order.id` and `product.name` could be used. `order.id` represents each unique order and `product.name` represents product name that is associated with this order.id. 

First of all, `order.id` and `product.name` will be extracted from `store_data` and then it will be stored in a new data frame named `prod_df`. 

Secondly, product qualification would be removed to find more match of products. If a product's name is "Kraft Clasp Envelope, Set of 50", then the "Set of 50" would be removed to increase the product count. 

Thirdly, using `dplyr` library, product names have been concatenated based on order.id. And afterward, it was saved as a new file named `transactions.csv`. 

```{r warning=FALSE}
#data frame with order.id and product.name have been created
sd_ord_prod <- data.frame(store_data$Order.ID, store_data$Product.Name)
colnames(sd_ord_prod) <- c("order.id", "product.name")

#product description has been removed
sd_ord_prod$product.name <- gsub(",.*", "", sd_ord_prod$product.name)

#converted the columns into factors
sd_ord_prod$order.id <- as.factor(sd_ord_prod$order.id)
sd_ord_prod$product.name <- as.factor(sd_ord_prod$product.name)

#using ddply library, based on user ID, the product names concatenated. 
  #head(sd_ord_prod)
transaction_data <- ddply(sd_ord_prod, c("order.id"), function(df1)paste(df1$product.name, collapse= ", "))

#afterwards the file has been saved as csv format
write.csv(transaction_data, "D:/Education/JKU 2023/Summer/Data Mining/market_basket_analysis/transactions.csv", quote = FALSE, row.names = FALSE)

```


In order to run association algorithm among categories, order.id and sub.category could be used. order.id represents each unique order and sub.category represents category name that is associated with this order.id. 

First of all, order.id and sub.category will be extracted from `store_data` and then it will be stored in a new data frame named `prod_df`. 

Secondly, using `dplyr` library, category names have been concatenated based on order.id. And afterward, it was saved as a new file named `cat_transaction.csv`. 


```{r warning=FALSE}

sd_order_cat <- data.frame(store_data$Order.ID, store_data$Sub.Category)
colnames(sd_order_cat) <- c("order.id", "sub.category")

sd_order_cat$order.id <- as.factor(sd_order_cat$order.id)
sd_order_cat$sub.category <- as.factor(sd_order_cat$sub.category)
#head(sd_ord_prod)
cat_transaction_data <- ddply(sd_order_cat, c("order.id"), function(df1)paste(df1$sub.category, collapse= ", "))

write.csv(cat_transaction_data, "D:/Education/JKU 2023/Summer/Data Mining/market_basket_analysis/cat_transactions.csv", quote = FALSE, row.names = FALSE)
```


## b. Mining Association Rules

In the following the, the `transaction.csv` is being converted to transactions data so that it can be run using the apriori algorithm 

```{r message=FALSE, warning=FALSE}
tr <- read.transactions("D:/Education/JKU 2023/Summer/Data Mining/market_basket_analysis/transactions.csv", format ="basket", sep = ",")
#summary(tr)
```


Here it shows the list of top 20 product items by their frequency. 

```{r warning=FALSE}
itemFrequencyPlot(tr, topN = 20, type ="absolute", col = brewer.pal(8, 'Pastel2'), main ="Absolute Item Frequency Plot")
```

In the following an association rule algorithm has been ran with 0.0001 support and .1 confidence. 0.0001 support means that customers bought an item set at least 0.01% of the whole transactions. In this case, this is not a significant. However, if support number is increased more than 0.01%, then the association rules do not reutrn any association. 

```{r warning=FALSE}
prod_association_rules <- apriori(tr, parameter = list(supp =0.0001, conf = 0.1, maxlen = 10))


```

* The apriori algorithm produced 12 item set.They have a very low count and only the top 5 has a relatively high confidence level (0.42-0.75). 

* **Staples** are often bought together with items such as **chairs, vaccume, xerox, envelop, legal cart, post binders, table.** Many of the items that are bought together with staples belong to office supply category. 

* Also, printer and Eldon box and Index Tab are bought together. Interstingly toaster and 3 hole punch are bought together. 

* As mentioned, these associations are not strong because their support count is low even for some of them the confidence level is higher. 

* Lift is a measure that shows how likely that bundle purchase will occur. More than 1 shows that it is likely. All the cases have more than 1 lift. Therefore, the are likely to occur. 


```{r warning=FALSE}
inspect(prod_association_rules)
```

* Kulczynski coefficient shows the strength of association between two sets of binary data. 1 means that they have perfect association or conditional probability where 0 means that they are independent. 

* From the analysis, it seems that they have a very weak Kulczynski coefficient. 

```{r warning=FALSE}
interestMeasure(prod_association_rules, measure = "Kulczynski", transactions = transactions)
```

```{r warning=FALSE}
plot(prod_association_rules,method="two-key plot")

```

* Since it is quite a low number, it means that dataset is quite imbalanced

```{r warning=FALSE}
store_df <- as.data.frame(store_data)
#head(store_df)
imbalanceRatio(store_df , classAttr = "Product.Name")
```


```{r warning=FALSE}
cat_tr <- read.transactions("D:/Education/JKU 2023/Summer/Data Mining/market_basket_analysis/cat_transactions.csv", format ="basket", sep = ",")
```

```{r warning=FALSE}
cat_association_rules <- apriori(cat_tr, parameter = list(supp =0.0006, conf = 0.5))
```
The following table shows category sets and their frequency with confidence level. 0.6% of all the transactions has Binders category in association with  	{Fasteners, Labels, Machines}, {Appliances, Fasteners, Phones}, {Appliances, Fasteners, Paper}, {Appliances, Art, Fasteners} etc. 

```{r warning=FALSE}
inspect(cat_association_rules)
```

```{r warning=FALSE}
plot(cat_association_rules,method="two-key plot")
```



## c. Mining Multilevel Association Rule

```{r warning=FALSE}
library(arules)
df <- read.csv("store.csv")

n_df <- data.frame(df$Order.ID, df$Sub.Category, df$Product.Name)
colnames(n_df) <- c("ord.id", "cat", "prod")

n_df$prod <- gsub(",.*", "", n_df$prod)

n_df_agg <- aggregate(prod ~ ord.id + cat, data = n_df, toString)

n_df_agg <- n_df_agg[,-1]
#multilevel <- addAggregate(n_df_agg, "cat")


#
#itemInfo(trans)

head(n_df_agg)
fin <-  paste(n_df_agg$cat, n_df_agg$prod, sep = ", ")
#head(fin)
#write.csv(fin, "D:/Education/JKU 2023/Summer/Data Mining/market_basket_analysis/test.csv", quote = FALSE, row.names = FALSE)

trans <- read.transactions("D:/Education/JKU 2023/Summer/Data Mining/market_basket_analysis/test.csv", format ="basket", sep = ",")

rules <- apriori(trans, parameter = list(supp = 0.001, conf = 0.5))
inspect(rules)

```

```{r warning=FALSE}
plot(rules,method="two-key plot")
```




# Task 2: Health Data

## a. Data Preparation

### a.1 Familiarize yourself with dataset for product orders

```{r warning=FALSE}
health_df <- read.csv("D:/Education/JKU 2023/Summer/Data Mining/market_basket_analysis/health.csv")


```

### a.2 Data Preparation

```{r warning=FALSE}
library(dplyr)
#creating symptoms data frame
sdf <- filter(health_df, type == "S")
head(sdf, 10)
```

```{r warning=FALSE}
#conditions data frame
cdf <- filter(health_df, type == "C")
head(cdf,12)
```

```{r warning=FALSE}
#treatment dataframe
tdf <- filter(health_df, type == "T")
head(tdf, 12)
```

```{r warning=FALSE}
# symptoms aggregation
sdf$reportid <- as.character(sdf$reportid)

```

Aggregating all the symptoms based on report id and saving the csv
```{r warning=FALSE}
sdf_agg <- aggregate(name ~ reportid, data = sdf, toString)

sdf_agg <- sdf_agg[,-1]
write.csv(sdf_agg, "D:/Education/JKU 2023/Summer/Data Mining/market_basket_analysis/s.csv", quote = FALSE, row.names = FALSE)

```

Aggregating all the conditions based on report id and saving the csv

```{r warning=FALSE}
cdf_agg <- aggregate(name ~ reportid, data = cdf, toString)

cdf_agg <- cdf_agg[,-1]
write.csv(cdf_agg, "D:/Education/JKU 2023/Summer/Data Mining/market_basket_analysis/c.csv", quote = FALSE, row.names = FALSE)

```

Aggregating all the treatments based on report id and saving the csv

```{r warning=FALSE}
tdf_agg <- aggregate(name ~ reportid, data = tdf, toString)
tdf_agg <- tdf_agg[,-1]
write.csv(tdf_agg, "D:/Education/JKU 2023/Summer/Data Mining/market_basket_analysis/t.csv", quote = FALSE, row.names = FALSE)
```

#### Association alalysis for symptoms

```{r warning=FALSE, message= FALSE }
library(arules)
s_trans <- read.transactions("D:/Education/JKU 2023/Summer/Data Mining/market_basket_analysis/s.csv", format ="basket", sep = ",")

```

* After filtering top 20 associations by lift, we can observe the following
  * Shoulder pain and hip pain have the most high lift (7.055) which means they are 7 times more likely to occur
  * Neck pain and pelvic pain are also associated strongly
  * Constipation has a strong association with adominal pain and diarrhea which is intuitive to understand
  * And more!
  
```{r warning=FALSE}
srules <- apriori(s_trans, parameter = list(supp = .03, conf = .5))
srules_sort <- sort(srules, by = "lift", decreasing = TRUE)
inspect(head(srules_sort, n = 20))

```

Kulczynski Coefficient

```{r warning = FALSE}
interestMeasure(srules_sort, measure = "Kulczynski", transactions = transactions)

```



```{r warning = FALSE}
interestMeasure(srules_sort, measure = "Kulczynski", transactions = transactions)

```

```{r echo=FALSE, warning=FALSE}
plot(srules_sort, method="two-key plot")
```

#### Association alalysis for conditions

```{r warning=FALSE, message= FALSE}

c_trans <- read.transactions("D:/Education/JKU 2023/Summer/Data Mining/market_basket_analysis/c.csv", format ="basket", sep = ",")

```

* Some intersting associations are given in the following
  * anxiety and chronic fatigue and depression 
  * irritable bowel syndrome and chronic fatigue
  * idiopathic hyperinsonmia and deepresion 

```{r warning=FALSE}
crules <- apriori(c_trans, parameter = list(supp = .02, conf = .5))
crules_sort <- sort(crules, by = "lift", decreasing = TRUE)
inspect(head(crules_sort, n = 20))

```
Kulczynski Coefficient

```{r warning = FALSE}
interestMeasure(crules_sort, measure = "Kulczynski", transactions = transactions)

```

```{r echo= FALSE, warning=FALSE}
plot(crules, method="two-key plot")
```

#### Association alalysis for treatments

```{r warning=FALSE, message= FALSE}

t_trans <- read.transactions("D:/Education/JKU 2023/Summer/Data Mining/market_basket_analysis/t.csv", format ="basket", sep = ",")

```

```{r, warning=FALSE}
trules <- apriori(t_trans, parameter = list(supp = .02, conf = .5))

trules_sort <- sort(trules, by = "lift", decreasing = TRUE)
inspect(head(trules_sort, n = 20))
```

Kulczynski Coefficient

```{r warning = FALSE}
interestMeasure(trules_sort, measure = "Kulczynski", transactions = transactions)

```

```{r echo= FALSE, warning=FALSE}
plot(trules, method="two-key plot")
```


Kulczynski Coefficient

```{r warning = FALSE}
interestMeasure(prod_association_rules, measure = "Kulczynski", transactions = transactions)

```

Imbalance Ratio
The dataset has a very low imbalance

```{r}
#head(store_df)
imbalanceRatio(health_df , classAttr = "type")
```
